### 2015-06-17 chapters/ipa_meets_unicode.tex
update IPA homoglyphs section

Some other homoglyphs related to encoding IPA in the Unicode Standard are:

  \item The uses of the apostrophe has led to long discussions on the Unicode
     Standard email list. An English keyboard inputs <{\fontspec{Monaco}'}> \textsc{apostrophe}
     at \uni{0027}, although the preferred Unicode apostrophe is the <\ ' >
     \textsc{right single quotation mark} at \uni{2019}.
  \item The glottal stop/glottalization/ejective marker is another completely
      different character <{\large ʼ}>, the \textsc{modifier letter apostrophe} at
      \uni{02BC}, but unfortunately looks mostly highly similar to
      \uni{2019}. 
  \item Another problem is the <ˁ> \textsc{modifier letter reversed
     glottal stop} at \uni{02C1} vs.\@ the <ˤ> \textsc{modifier
     letter small reversed glottal stop} at \uni{02E4}. Both 
     appear in various resources representing phonetic data online. This is
     thus a clear example for which the Unicode Standard does not solve the
     linguistic standardization problem.
  \item even linguists are unlikely to distinguish between the <ə>
     \textsc{latin small letter schwa} at code point \uni{0259} and <ǝ>
     \textsc{latin small letter turned e} at \uni{01DD}.
  \item There is at least one case in which the character name assigned by the
     Unicode Consortium does not match the IPA's description. In the Unicode
     Standard the <ǃ> at \uni{01C3} is labeled \textsc{latin letter retroflex
     click}, but in IPA that glyph is used for an alveolar or postalveolar
     click (not retroflex). This naming is probably best seen as a simple
     error in the Unicode Standard. This character is of course often simply
     typed as <!> \textsc{exclamation mark} at \uni{0021}.

@@

Conversely,
non-linguists are unlikely to distinguish any semantic difference between an
open back unrounded vowel <ɑ> \textsc{latin small letter alpha} at
\uni{0251}, and the open front unrounded vowel <a> \textsc{latin small letter
a} at \uni{0061}. But even among linguists this distinction leads to problems.
For example, as pointed out by \citet{Mielke2009}, there is a problem stemming
from the fact that about 75\% of languages are reported to have a five-vowel
system \citep{Maddieson1984}. Historically, linguistic descriptions tend not to
include precise audio recording and measurements of formants, so this may lead
one to ask if the many <a> characters that are used in phonological description
reflects a transcriptional bias. The common use of <a> in transcriptions
could be in part due to the ease of typing the letter on an English keyboard (or
for older descriptions, the typewriter). We found it to be exceedingly rare that
a linguist uses <ɑ> for a low back unrounded vowel.\footnote{One example is
\citet[75]{Vidal2001a}, in which the author states: ``The definition of Pilagá
/a/ as [+back] results from its behavior in certain phonological contexts.
For instance, uvular and pharyngeal consonants only occur around /a/ and /o/.
Hence, the characterization of /a/ and /o/ as a natural class of (i.e.,
[+back] vowels), as opposed to /i/ and /e/.''} They simply use <a> as
long as there is no opposition to <ɑ>.\footnote{See Thomason's Language Log
post, ``Why I don't love the International Phonetic Alphabet'' at:
\url{http://itre.cis.upenn.edu/~myl/languagelog/archives/005287.html}.}

>>>>

\noindent Some problematic homoglyphs related to encoding IPA in the Unicode Standard are 
the following.
\begin{itemize}

  \item The uses of the apostrophe has led to long discussions on the Unicode
     Standard email list. An English keyboard inputs <{\fontspec{Monaco}'}> \textsc{apostrophe}
     at \uni{0027}, although the preferred Unicode apostrophe is the <\ ' >
     \textsc{right single quotation mark} at \uni{2019}.\footnote{Note that many word 
     processors (like Microsoft Word) by default will replace straight quotes by 
     curly quotes, depending on the whitespace around it. Such a replacement is 
     a real character replacement, meaning that another Unicode code point is 
     inserted.} However, the glottal stop/glottalization/ejective marker is yet 
     another completely different character <{\large ʼ}>, the 
     \textsc{modifier letter apostrophe} at \uni{02BC}, which unfortunately looks 
     mostly extremely similar to \uni{2019}. 
  \item Another problem is the <ˁ> \textsc{modifier letter reversed
     glottal stop} at \uni{02C1} vs.\@ the <ˤ> \textsc{modifier
     letter small reversed glottal stop} at \uni{02E4}. Both 
     appear in various resources representing phonetic data online. This is
     thus a clear example for which the Unicode Standard does not solve the
     linguistic standardization problem.
  \item even linguists are unlikely to distinguish between the <ə>
     \textsc{latin small letter schwa} at code point \uni{0259} and <ǝ>
     \textsc{latin small letter turned e} at \uni{01DD}.
  \item  The alveolar click <ǃ> at \uni{01C3} is of course often simply
     typed as <!> \textsc{exclamation mark} at \uni{0021}.\footnote{In the Unicode
     Standard the <ǃ> at \unif{01C3} is labeled \textsc{latin letter retroflex
     click}, but in IPA that glyph is used for an alveolar or postalveolar
     click (not retroflex). This naming is probably best seen as an 
     error in the Unicode Standard.}

@@

Conversely, non-linguists are unlikely to distinguish any semantic difference
between an open back unrounded vowel <ɑ> \textsc{latin small letter alpha} at
\uni{0251}, and the open front unrounded vowel <a> \textsc{latin small letter a}
at \uni{0061}, basically treating them as homoglyphs, although they are
different phonetic symbols.

But even among linguists this distinction leads to problems. For example, as
pointed out by \citet{Mielke2009}, there is a problem stemming from the fact
that about 75\% of languages are reported to have a five-vowel system
\citep{Maddieson1984}. Historically, linguistic descriptions tend not to include
precise audio recording and measurements of formants, so this may lead one to
ask if the many <a> characters that are used in phonological description
reflects a transcriptional bias. The common use of <a> in transcriptions could
be in part due to the ease of typing the letter on an English keyboard (or for
older descriptions, the typewriter). We found it to be exceedingly rare that a
linguist uses <ɑ> for a low back unrounded vowel.\footnote{One example is
\citet[75]{Vidal2001a}, in which the author states: ``The definition of Pilagá
/a/ as [+back] results from its behavior in certain phonological contexts. For
instance, uvular and pharyngeal consonants only occur around /a/ and /o/. Hence,
the characterization of /a/ and /o/ as a natural class of (i.e., [+back]
vowels), as opposed to /i/ and /e/.''} They simply use <a> as long as there is
no opposition to <ɑ>.

%\footnote{See Thomason's Language Log post, ``Why I don't love the International Phonetic Alphabet'' at:\url{http://itre.cis.upenn.edu/~myl/languagelog/archives/005287.html}.}
